---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(dev = c('png', 'pdf'), dpi = 300, dev.args = list(bg = 'transparent'), eval = FALSE)
```

# the iris dataset (not belonging to the tidyverse)

```{r}
iris

iris %>% 
  split(iris$Species) %>% 
  fsf::xview(list_collapse = FALSE) # use '::' to access installed R-packages (without 'library')
```

# manipulate data

```{r}
iris_backup <- iris # always a good idea to create "backups"

# introduceing some random NA values
iris$Sepal.Length[sample(1:150, 30)] <- NA
iris$Petal.Width[sample(1:150, 13)] <- NA
iris$Petal.Length[sample(1:150, 1)] <- NA
```

# resources

- [tidyverse intro, topic on tidy data](https://tidyverse-intro.github.io/4-tidy.html#tidy)
- [tidyverse packages logo](https://cdn.analyticsvidhya.com/wp-content/uploads/2019/05/ggplot_hive-768x392.jpg)
- [advanced](https://dplyr.tidyverse.org/articles/programming.html)

# load the tidyverse (bunch of packages) as you would load a normal package

```{r}
library(tidyverse)
```

# the maggritr pipe %>% 

```{r}
# base R:
str(iris) # same as
str(object = iris)

# maggritr workflow:
iris %>% str # same as...
iris %>% str() # implicit pipe
iris %>% str(object = .) # explicit pipe
```

- with or without dot '.'
    - no difference when placeholder should specify the **first** function argument (dot assumed for the very first argument --> all tidyverse functions are coded with that in mind = consistent syntax)
    - `.` is needed to pass data to second, third, ... argument 
    
```{r}
iris %>% str(object = .) # explicitly passing an object to the first function argument (not necessary)

lm(iris$Sepal.Length ~ iris$Sepal.Width)
lm(Sepal.Length ~ Sepal.Width, data = iris)

iris %>% lm(Sepal.Length ~ Sepal.Width, data = .) # now '.' is necessary because object is passed to argument 'data', which is not the first one
```

- brackets vs pipes *or* base vs "tidyverse"

```{r}
summary(lm(Sepal.Length ~ Sepal.Width, data = iris)) # base R
iris %>% 
  lm(Sepal.Length ~ Sepal.Width, data = .) %>% 
  summary # using maggritr pipe
```

# functions

arrange  
filter  
mutate  
summarize  
group_by  
  
select  
sample_n
slice  
rename  

# arrange

```{r}
iris %>% 
  arrange(desc(Sepal.Length), Sepal.Width) %>% 
  head()
```

# filter NA values

```{r}
iris # simply look at the data and search for NA... bad idea 
summary(is.na(iris)) # better, but only colwise

iris %>% 
  is.na() %>% 
  colSums()

iris %>% 
  is.na() %>% 
  rowSums() %>% 
  table()

visdat::vis_miss(iris) # offtopic, but handy function

iris %>% filter(!is.na(Sepal.Length)) # filter na values in specific column

iris %>% filter_all(all_vars(!is.na(.))) # filter all !na

iris %>% # filter rows in columns beginning with "Petal", that include NA values
  filter_at(vars(starts_with("Petal")), any_vars(is.na(.)))
```

# mutate

- create new columns (same nrow as data frame)

```{r}
miris <- iris %>% 
  mutate(long_sepals = Sepal.Length > mean(Sepal.Length),
         wide_sepals = Sepal.Width > mean(Sepal.Width)) %>% 
  mutate(big_sepals = long_sepals + wide_sepals) 

# always check the output!
miris %>% str() # not as intended.... what happened?

miris <- iris %>% 
  filter_all(all_vars(!is.na(.))) %>%  #!
  mutate(long_sepals = Sepal.Length > mean(Sepal.Length),
         wide_sepals = Sepal.Width > mean(Sepal.Width)) %>% 
  mutate(big_sepals = long_sepals + wide_sepals) 

# check again
miris %>% str() # reminder, that you always should be aware of NA values (e.g. you cannot calculate the mean of a vector containing NAs)
```

# summarize

- aggregate columns (**reduces dimensionality!**)

```{r}
miris %>% 
  summarize(MEAN_SW = mean(Sepal.Width),
            MEAN_SL = mean(Sepal.Length),
            SD_SL = sd(Sepal.Length),
            MEDIAN_SL = median(Sepal.Length)) # ...
```

# group_by

- do something for each group

```{r}
miris %>% 
  count(big_sepals) #   count(long_sepals, wide_sepals)

miris %>% 
  group_by(Species) %>% 
  count(big_sepals)
```

# slice and select

- non-random subsetting (compare with sample)
    - rows: slice
    - cols: select

```{r}
# "top n by group" example
iris %>% 
  group_by(Species) %>% 
  arrange(desc(Sepal.Width)) %>%
  slice(1:3) %>% 
  ungroup() %>% 
  select(Species, Sepal.Width)
```




